Temperature and wave exposure as drivers of morphology in two kelp species

### 2019-07-24 morphology_paper.Rmd
upadted rmd file, renamed. Script reduced. test outputs added.

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  word_document: 
    fig_caption: yes
    reference_docx: frontiers_template.docx
  pdf_document: 
    fig_caption: yes
    fig_crop: no
    fig_height: 7
    fig_width: 7
  html_notebook: default
papersize: A4
tables: yes
fontsize: 10pt
---

@@

```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(gridExtra)
```

@@

```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}

### Plots for temperature variables

positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")

# Annual temp
temp_plot_A <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) + 
  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("A") +
  theme(plot.title = element_text(hjust = 0)) +
  theme(legend.position = "none") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
# Summer temp
temp_plot_B <- ggplot(data = temp_clim, aes(x = site, y = Feb_mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) +
  geom_errorbar(aes(x = site, ymin = Feb_mean_temp - Feb_sd_temp, ymax = Feb_mean_temp + Feb_sd_temp   )) + 
  geom_point(data = temp_clim, aes(x = site, y = Feb_max_temp, colour = "blue")) + 
  geom_point(data = temp_clim, aes(x = site, y = Feb_min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("B") +
  theme(legend.position = "none") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

# Winter temp
temp_plot_C <- ggplot(data = temp_clim, aes(x = site, y = Aug_mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) + 
  geom_errorbar(aes(x = site, ymin = Aug_mean_temp - Aug_sd_temp, ymax = Aug_mean_temp + Aug_sd_temp   )) + 
  geom_point(data = temp_clim, aes(x = site, y = Aug_max_temp, colour = "blue")) + 
  geom_point(data = temp_clim, aes(x = site, y = Aug_min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("C") +
  theme(legend.position = "none") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) 
```

```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots.", fig.height=8, fig.width=8}

temp_plot_all <- grid.arrange(temp_plot_A, temp_plot_B,temp_plot_C, ncol = 1, nrow = 3 )

ggsave("temperature_plot.png", plot = temp_plot_all, path = "figures/", dpi = 300, width = 5, height = 5, limitsize = TRUE, scale = 2)

@@

```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

## Data wrangling and preparation

# Wave data

# Load the loading functions
source("func/load.wave.R")

# Then load the data
fb_model <- load.model("data/wave_data/FB")
tb_model <- load.model("data/wave_data/TB")
he_model <- load.model("data/wave_data/HE")
sh_model <- load.model("data/wave_data/SH")

@@

 # Calculate the necessary stats
# Monthly clims
wave_feb_aug <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
wave_ann <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean), na.rm = T) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# Subset directions out
dirs <- subset(wave_model, select = c(1, 2, 5, 6))

# Seasons
dirs_feb_aug <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual
dirs_ann <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>% 
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# subset useless columns
dirs_feb_aug <- subset(dirs_feb_aug, select = c(1, 2, 6, 10))
dirs_ann <- subset(dirs_ann, select = c(1, 2, 6, 10))

# combine all direction metrics
dirs_clim <- rbind(dirs_feb_aug, dirs_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         3:4) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# Combine
# NB: This only uses the wave data from the 7 metre contour
wave_clim <- rbind(wave_feb_aug, wave_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         hs_mean:spw_sd) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# combine all wave metrics
wave_clim <- cbind(wave_clim, dirs_clim)
# remove all useless direction metrics from original wave_clim
wave_clim <- subset(wave_clim, select = -c(2,3,4,5,12,13,14,15,22,23,24,25,32,33,40))
```

```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}

### Plots for wave variables
#Annual wave plots

wave_plot_A <- ggplot(data = wave_clim, aes(x = site, y = Ann_hs_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Ann_hs_mean - Ann_hs_sd, ymax = Ann_hs_mean + Ann_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("A") + 
  theme(plot.title = element_text(hjust = 0)) + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

wave_plot_B <- ggplot(data = wave_clim, aes(x = site, y = Ann_tp_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Ann_tp_mean - Ann_tp_sd, ymax = Ann_tp_mean + Ann_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("B") + 
  theme(plot.title = element_text(hjust = 0)) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip()

#Summer wave plots
wave_plot_C <- ggplot(data = wave_clim, aes(x = site, y = Feb_hs_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Feb_hs_mean - Feb_hs_sd, ymax = Feb_hs_mean + Feb_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("C") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

wave_plot_D <- ggplot(data = wave_clim, aes(x = site, y = Feb_tp_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Feb_tp_mean - Feb_tp_sd, ymax = Feb_tp_mean + Feb_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
    ggtitle("D") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

# Winter summary temp
wave_plot_E <- ggplot(data = wave_clim, aes(x = site, y = Aug_hs_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Aug_hs_mean - Aug_hs_sd, ymax = Aug_hs_mean + Aug_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  ggtitle("E") + 
  theme(legend.position = "none") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

# Winter temperature range
wave_plot_F <- ggplot(data = wave_clim, aes(x = site, y = Aug_tp_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Aug_tp_mean - Aug_tp_sd, ymax = Aug_tp_mean + Aug_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("F") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) 
```

```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers.', fig.height=8, fig.width=8}

wave_plot_all <- grid.arrange(wave_plot_A, wave_plot_B, wave_plot_C, wave_plot_D, wave_plot_E, wave_plot_F, ncol = 2, nrow = 3)

ggsave("wave_plot.png", plot = wave_plot_all, path = "figures/", dpi = 300, width = 6, height = 6, limitsize = TRUE, scale = 2)

@@

@@

#wave exposure and lamanaria
lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")

@@

```{r laminaria plots, echo=FALSE, fig.cap='Boxplots of *L. pallida morphological characteristics group by temperature regime and wave exposure category.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}

ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
```

@@

library(ggplot2)

@@

```{r deep ecklonia  plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature regime and wave exposure.'}
ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8, EW9, EW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

@@

@@

@@

@@

```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime and wave exposure.'}

ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")

@@

>>>>

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  pdf_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 7
    fig_width: 7
  html_notebook: default
  html_document:
    df_print: paged
  word_document:
    reference_docx: frontiers_template.docx
papersize: A4
tables: yes
fontsize: 10pt
---

@@

```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(ggplot2)
library(ggord)
library(vegan)
```

@@

```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}
### Plots for temperature variables 
positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")

# Annual temp
temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) + 
  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  theme(plot.title = element_text(hjust = 0)) +
  theme(legend.position = "none") + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

@@

```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap='Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
temp_plot
```

```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
## Data wrangling and preparation
# Wave data 
# Load the loading functions
source("func/load.wave.R")
# Then load the data
fb_model <- load.model("data/wave_data/FB")
tb_model <- load.model("data/wave_data/TB")
he_model <- load.model("data/wave_data/HE")
sh_model <- load.model("data/wave_data/SH")

@@

 # Calculate the necessary stats
# Monthly clims
wave_feb_aug <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
wave_ann <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean), na.rm = T) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything()) %>% 
  group_by(site) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

wave_ann <- wave_ann[-c(9,15,20,22), -c(2,23)]

#remove useless coloumns
wave_ann <- wave_ann %>% 
  select(site, hs_mean, hs_min, hs_max, hs_sd, tp_mean, tp_min, tp_max, tp_sd) 
```

```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}
### Plots for wave variables
#Annual wave plots

wave_plot_A <- ggplot(data = wave_ann, aes(x = site, y = hs_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = hs_mean - hs_sd, ymax = hs_mean + hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  theme(plot.title = element_text(hjust = 0)) + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

wave_plot_B <- ggplot(data = wave_ann, aes(x = site, y = tp_mean)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = tp_mean - tp_sd, ymax = tp_mean + tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  theme(plot.title = element_text(hjust = 0)) + 
  coord_flip() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())  
```

```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers.'}
ggarrange(wave_plot_A, wave_plot_B, ncol = 2, nrow = 1)
```

@@

  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 

@@

lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))#wave exposure levels for lamanaria
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")#custom positions

@@

```{r laminaria plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime and wave exposure category.', message=FALSE, warning=FALSE, include=FALSE}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
```

@@

@@

```{r deep ecklonia plot, message=FALSE, warning= FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature regime and wave exposure.'}
ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10,EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
```

@@

  ylim(0, 20)

@@

  ylim(0, 30)

@@

  guides(fill=guide_legend(title="Wave exposure level")) + 
  ylim(0,25)

@@

```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime and wave exposure.'}
ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
```

```{r echo=FALSE}
#calculate mean values for lamanaria morphology
lam_mean <- lam_df %>% 
  subset(select = -c(1,2)) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean), na.rm = T) %>% 
  ungroup()

# merge temp and wave data to create environmental data-frame
abio_clim <- merge(temp_ann, wave_ann, by = "site")

#remove date column
abio_clim <- subset(abio_clim, select = -c(2))

# standardise measurements
abio_clim <- abio_clim %>%
  subset(select = -c(1)) %>% 
  decostand(method = "standardize")

# run RDA
RDA_all_eck <- rda(eck_bio_deep ~ ., data = eck_abio_deep)

# Plot RDA
plot(RDA_all_eck, scaling  = 2)
```

@@

# Conclusion

Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables. We confirm that wave variables are the main driver of kelp morphological characteristics in both *E. maxima *and *L. pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E. maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.

# Conflict of Interest Statement

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest

# Author Contributions

RC conceptualised the scope of the research reported in this paper, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing. AS and CR provided guidance in terms of  writing, statistical analyses, and the final draft of the paper, as well some writing in the relevant sections. RS aided in fieldwork and statistical analyses for the temperature data. TP performed the initial analyses and interpretation of the data.

# Funding

The research was funded by the South African National Research Foundation ([http://www.nrf.ac.za](http://www.nrf.ac.za)) programme "Thermal characteristics of the South African nearshore:implications for biodiversity" (CPRR14072378735).

Aside from funding provided, the funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.

# Acknowledgments

Thank you to AS and CR for their guidance, and the honours class of 2014-2015 that aided in fieldwork. Carla Louise for making the GIS plot etc.
